# Part A: Loading and Processing the data

## A1: Code needed to load the data

```{r, echo=TRUE}

#Loading and preprocessing the data
act<-read.csv("C:/Users/Dan_Wolf/Documents/DataScience/JHU5_RR/activity.csv",na.strings="NA")
head(act)
summary(act)

```

## A2: Preprocess the data 
Make interval a factor
```{r, echo=TRUE}
#make interval a factor
act$interval <- factor(act$interval)
#install.packages("timeDate")
library(timeDate)
#Create a column that indicates of a day is a weekday or weekend
act$dayType<-ifelse(isWeekday(act$date, wday=1:5),"Weekday","Weekend")
# check 
table(act$dayType)


```


# Part B: What is mean total number of steps taken per day?

## B1: Calculate Total number of steps taken per day 

```{r, echo=TRUE,results="hide"}

#install.packages("plyr")
library(plyr)
ddply(act,.(date),summarize, "mean"=mean(steps,na.rm=TRUE), "median" = median(steps,na.rm=TRUE),"sum"=sum(steps,na.rm=TRUE))

```

## B2: Histogram of total steps taken each day

```{r, echo=TRUE}
StepsPerDay<-ddply(act,.(date),summarize,"sum"=sum(steps,na.rm=TRUE), "mean"=mean(steps,na.rm=TRUE), "median" = median(steps,na.rm=TRUE))
#install.packages("ggplot2")
library(ggplot2)
hist(StepsPerDay$sum,breaks=10,main="Distribution of Total Steps Per Day")
```

## B3: Mean, median and total number (Sum) of steps taken each day
```{r, echo=TRUE}
print (StepsPerDay)
```


# Part C: What is average daily activity pattern? 

## C1: Plot 5 min interval vs average number of steps taken averaged across all days


```{r, echo=TRUE}

meanStepsPerInterval<-ddply(act,.(interval),summarize,"mean"=mean(steps,na.rm=TRUE))
plot(x<-as.numeric(paste(meanStepsPerInterval$interval)),y<-meanStepsPerInterval$mean,type="l",xlab="Interval Start Time",ylab="Average Total Steps per 5min Interval", main="Average Total Steps Per 5 Min Interval")

```


## C2: Which 5 min interval across all days in the data set contains the max number of steps? 

```{r, echo=TRUE}
# find max sum
sumStepsPerInterval<-ddply(act,.(interval),summarize,"sum"=sum(steps,na.rm=TRUE))
sumStepsPerInterval$sum<-as.numeric(sumStepsPerInterval$sum)
sumStepsPerIntervalOrdered <- sumStepsPerInterval[order(sumStepsPerInterval$sum,decreasing=TRUE),]
maxInterval<-as.numeric(as.character(sumStepsPerIntervalOrdered$interval[1]))
maxSum<- as.numeric(sumStepsPerIntervalOrdered$sum[1])
print (maxSum)
print (maxInterval)
```

The 8:35am to 8:40am interval contains the max number of steps across all days at 10927 steps.


# Part D: Imputing missing values: 

## D1: Total number of rows with steps values missing

```{r, echo=TRUE}
missing <- sum(is.na(act$steps))
print (missing)
```

There are 2304 rows with missing values


## D2: Strategy for imputing missing values: I'll use mean for 5 min intervals.  This requires merging the 5min mean file from above with the act data frame. 

## D3: Create new data set

This is tricky - need to merge two data sets using 'interval' as the match key. 
Then create new steps_imputed column (I wanted to save the old, unimputed value)

```{r, echo=TRUE}
act_merged<-merge(act,meanStepsPerInterval,by="interval")

for (row in act_merged) 
     { 
       act_merged$steps_imputed<-ifelse(is.na(act_merged$steps),act_merged$mean,act_merged$steps)
     }

```

Then create a new data frame with mean, median and sum

```{r,echo=TRUE}
StepsPerDayImputed<-ddply(act_merged,.(date),summarize,"sum"=sum(steps_imputed,na.rm=TRUE), "mean"=mean(steps_imputed,na.rm=TRUE), "median" = median(steps_imputed,na.rm=TRUE))

```

## D4: Historgram of total numbr of steps taken each day after imputations

```{r,echo=TRUE}

hist(StepsPerDayImputed$sum,breaks=10,main="Distribution of Total Steps Per Day after Imputations")
```
```{r,echo=TRUE}
print (StepsPerDayImputed)
```

## D5: Mean, median and total number (Sum) of steps taken each day after imputation

```{r,echo=TRUE}
print (StepsPerDayImputed)
```

## D6: Conclusion: Do these values differ from B3?  

## Yes - the historgram is more centered, less extreme values.  What is impact: The Sums and means are less extreme.  


# Part E: Are there differences between activity patterns between weekdays and weekends? 

## E1: Create new factors 

This was done in section A2


## E2: Graph Panel Plot of weekday and weekend activity

```{r,echo=TRUE}
# convert interval to numeric and date to character

act_merged$interval<-as.numeric(paste(act_merged$interval))
act_merged$date<-as.character(paste(act_merged$date))
#install.packages("rshape2")
library(reshape2)
# use melt to create a long table for data manipulation
actmelt <- melt(act_merged,id=c("interval","dayType"),measure.vars=c("steps_imputed"))

# take mean of steps imputed across dayType and interval
actmelt <- dcast(actmelt, dayType + interval ~ variable,fun.aggregate=mean)

#Plot
library(ggplot2)
sp <- ggplot(actmelt, aes(x=interval, y=steps_imputed),ylab="Average Steps Per Interval",xlab="Start Time of Interval") + geom_line() + facet_grid(dayType~.)
sp
```

## E3: Conclusion

## Looks like my mother when I was little!  Get up really early - get the kids off to school weekdays, Go running or gym 8:30 to 9:30, get to work by 10. Weekends sleep in then go running 8:30 to 9:30.

