#Part A: Loading and Processing the data

## A1: Code needed to load the data

```{r, echo=TRUE}

#Loading and preprocessing the data
act<-read.csv("C:/Users/Dan_Wolf/Documents/DataScience/JHU5_RR/activity.csv",na.strings="NA")
head(act)
summary(act)

```
## A2: Preprocess the data: 
###     Make interval a factor
```{r, echo=TRUE}
#make interval a factor
act$interval <- factor(act$interval)
#install.packages("timeDate")
library(timeDate)
#Create a column that indicates of a day is a weekday or weekend
act$dayType<-ifelse(isWeekday(act$date, wday=1:5),"Weekday","Weekend")
# check 
table(act$dayType)


```

#Part B: What is mean total number of steps taken per day?


## B1: Calculate Total number of steps taken per day 

```{r, echo=TRUE,results="hide"}

#install.packages("plyr")
library(plyr)
ddply(act,.(date),summarize, "mean"=mean(steps,na.rm=TRUE), "median" = median(steps,na.rm=TRUE),"sum"=sum(steps,na.rm=TRUE))

```

## B2: Histogram of total steps taken each day

```{r, echo=TRUE}

StepsPerDay<-ddply(act,.(date),summarize,"sum"=sum(steps,na.rm=TRUE), "mean"=mean(steps,na.rm=TRUE), "median" = median(steps,na.rm=TRUE))
#install.packages("ggplot2")
library(ggplot2)
hist(StepsPerDay$sum,breaks=10)
```

## B3: Mean, median and total number (Sum) of steps taken each day
```{r, echo=TRUE}
print (StepsPerDay)
```



#Part C: What is average daily Activity pattern? 

## C1: Plot 5 min interval vs avg number of steps taken averaged across all days


```{r, echo=TRUE}
meanStepsPerInterval<-ddply(act,.(interval),summarize,"mean"=mean(steps,na.rm=TRUE))
plot(x<-as.numeric(paste(meanStepsPerInterval$interval)),y<-meanStepsPerInterval$mean,type="l")
```


## C2: Which 5 min interval across all days in the data set contains the max number of steps? 


# Part D: Imputing missing values: 

## D1: Total number of missing values:

## D2: Strategy for imputing missing values: Use mean for 5 min intervals.  This requires merging the 5min mean file from above with the act data frame. 

## D3: Create new data set

## D4: Historgram of total numbr of steps taken each day

## D5: Mean, median and total number (Sum) of steps taken each day

## D6: Conclusion: Do these values differ fron B3? _____  What is impact: ___


# Part E: Are there differences between activity patterns between weekdays and weekends? 

## E1: Create new factos 

## E2: Panel Plots

## E3: Conclusion